# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data 
```{r loaddata}
activity.raw = read.csv(unz("activity.zip", "activity.csv"))
activity = na.omit(activity.raw)
```



## What is mean total number of steps taken per day?
```{r histogram, fig.height=5, fig.width=5 }
library(plyr)
perday2 = ddply(activity, "date", transform, day=sum(steps) )
hist(perday2$day, main= "Frequency of Steps Taken Each Day", xlab="Number of Steps")
```

```{r  meanandmedian}
mean(perday2$day)
median(perday2$day)
```



## What is the average daily activity pattern?
```{r timeseries}
byinterval = ddply(activity, "interval", transform, meanint=mean(steps))
with(byinterval, plot(interval, meanint,  type="l" ))
```


### What is the interval that has the largest average number of steps taken?
```{r maxinterval}
byinterval[which.max(byinterval$meanint), "interval"]
```




## Imputing missing values  HISTOGRAM

### What are the total number of NA values from the orignal data set?
```{r}
sum(is.na(activity.raw))
```

### Fill in all of the missing values in the dataset
```{r}
### this is the avaerage number of  step taken per interval
mean.place = mean(activity$step)
### this replaces NA from the original file with the average
activity.raw[is.na(activity.raw)] = mean.place
activity.replace = activity.raw
```


### How does replacing the NA values affect mean total number of steps taken per day?
```{r histogramNA, fig.height=5, fig.width=5 }
perdayNA = ddply(activity.replace, "date", transform, day=sum(steps) )
hist(perdayNA$day, main= "Frequency of Steps Taken Each Day", xlab="Number of Steps", breaks=11)
```

### What is the difference between the mean and the median after the NA values are replaced?
```{r meanandmediandiff}
mean(perday2$day) - mean(perdayNA$day)
median(perday2$day) - median(perdayNA$day)
```

#### Using the average number of steps per 5 minute interval for the total dataset to replace NA did not change the average number of steps and decreased the median by 1.2 steps. 


## Are there differences in activity patterns between weekdays and weekends?
```{r}
activity$day = weekdays(as.Date(activity$date))
library(chron)
activity$weekend = is.weekend(as.Date(activity$date))
```

